Crop diversification in a game against nature

### 2020-01-20 risk-percption.Rmd
add state of nature

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

@@

||Drought Year|Normal Year|
|-------|-----|-----|----|
|Barley Yield|0.93|1.18|
|Wheat Yield|0|1.60|

@@

```{r}
wheat_dry <- 0
wheat_wet <- 1.6
barley_dry <- 0.93
barley_wet <- 1.18
```

@@

The expected yield is thus calculated as.
```{r}
expected_utility <- function(risk) {
  risk * (wheat_dry + barley_dry) + (1 - risk) * (wheat_wet + barley_wet) 
}
```

```{r}
expected_utility(0.2) # assuming a drought risk of 0.2
```

@@

tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
  gather(type, utility, dry_exp:wet_exp) %>%
  ggplot(aes(wheat_prop, utility)) +
  geom_hline(yintercept = nash_eq$utility, linetype = 2)+
  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = nash_eq, color = 'black', size = 3) +
  geom_text(data = nash_eq, 
            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

@@

```{r}
library(raster)
library(CDFt)
```

@@

sim_present <- brick('~/Downloads/b40.20th.track1.1deg.005.cam2.h0.PRECT.185001-200512.nc') %>%
  raster::extract(konya, df = TRUE) %>%
  dplyr::select(-ID) %>%
  gather(time, rainfall) %>%
  mutate(rainfall = rainfall * 2.628e+9,
         year = rep(1850:2005, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
         month = rep(1:12, times = 156)) %>%
  dplyr::select(-time) %>%
  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
  filter(between(water_year, 1932, 2005)) %>%
  group_by(water_year) %>%
  summarise(precip = sum(rainfall))

sim_past <- brick('~/Downloads/b40.mh6ka.1deg.003.cam2.h0.PRECT.080101-131612.nc') %>% 
  raster::extract(konya, df = TRUE) %>%
  dplyr::select(-ID) %>%
  gather(time, rainfall) %>%
  mutate(rainfall = rainfall * 2.628e+9,
         year = rep(1:516, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
         month = rep(1:12, times = 516)) %>%
   dplyr::select(-time) %>%
  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
  group_by(water_year) %>%
  summarise(precip = sum(rainfall))
```

```{r}
obs <- read_table('~/Downloads/pa17244.dat', 
           comment = "#", 
           col_names = c('year', 1:12), 
           na = c('-999.9', '-888.8')) %>%
  gather(month, precip, `1`:`12`) %>%
 # filter(!(month %in% c(4:10)) & year < 2006) %>%
  group_by(month) %>%
  mutate(precip = if_else(is.na(precip), mean(precip, na.rm = TRUE), precip)) %>%
  ungroup %>%
  mutate(month = as.numeric(month),
         water_year = if_else(month %in% 10:12, year + 1, year)) %>% 
    filter(between(water_year, 1932, 2005)) %>%
  group_by(water_year) %>%
  summarise(precip = sum(precip))
```

@@

obs_past < 300
```

@@

so we're imagining a yield of 1 for barley and 2 for wheat, with equal prportions of each. then in the next time step we'd revise the proportion to take away .25
```{r}
reallocate <- function(wheat_prop, risk) {
  wheat_prop + 1 * (1.18 * (1 - wheat_prop) - 1.6 * wheat_prop) * wheat_prop * (1 - wheat_prop)
}

@@

tibble(rainfall = (0:200) / 100,
       barley = calc_yield(rainfall, 'barley'),
       wheat = calc_yield(rainfall, 'wheat')) %>%
  gather(crop, yield, barley:wheat) %>%
  ggplot(aes(rainfall, yield, color = crop)) + 
  geom_line() +
  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
  geom_hline(yintercept = yield_max_barley * kcal_barley) +
  geom_hline(yintercept = 1016525 * 5, color ='red') +
  geom_vline(xintercept = 0.3)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(raster)
library(tidyverse)
library(CDFt)
```

@@

||Dry Year|Normal Year|Wet Year|
|-------|-----|-----|----|----|
|Barley Yield|0.93|1.18||
|Wheat Yield|0|1.60||

@@

```{r}
#wheat_dry <- 0
#wheat_wet <- 1.6
#barley_dry <- 0.93
#barley_wet <- 1.18

wheat_dry <- 30
wheat_normal <- 60
wheat_wet <- 70

barley_dry <- 50
barley_normal <- 50
barley_wet <- 40
```

@@

The expected yield is thus calculated as.
```{r}
expected_yield_wheat <- function(p_dry, p_wet) {
  p_dry * wheat_dry + p_wet * wheat_wet + (1 - p_wet - p_dry) * wheat_normal 
}

expected_yield_barley <- function(p_dry, p_wet) {
  p_dry * barley_dry + p_wet * barley_wet + (1 - p_wet - p_dry) * barley_normal 
}
```

```{r}
expected_yield_wheat(p_dry = 0.3, p_wet = 0.2)
expected_yield_barley(p_dry = 0.3, p_wet = 0.2)
```

The risk averse MAXIMIN pure strategy is to plant barley all the time, and the risk indifferent farmer plants wheat for its high yields

```{r}
yield_var_wheat <- function(p_dry, p_wet) {
  exp_wheat <- expected_yield_wheat(p_dry, p_wet)
  p_dry * (wheat_dry - exp_wheat) ^ 2 +
    p_wet * (wheat_wet - exp_wheat) ^ 2 +
    (1 - p_dry - p_wet) * (wheat_normal - exp_wheat) ^ 2
}

yield_var_barley <- function(p_dry, p_wet) {
  exp_barley <- expected_yield_barley(p_dry, p_wet)
  p_dry * (barley_dry - exp_barley) ^ 2 +
    p_wet * (barley_wet - exp_barley) ^ 2 +
    (1 - p_dry - p_wet) * (barley_normal - exp_barley) ^ 2
}

yield_var_wheat(0.3, 0.2)
yield_var_barley(0.3, 0.2)
```

```{r}
yield_mad_wheat <- function(p_dry, p_wet) {
  exp_wheat <- expected_yield_wheat(p_dry, p_wet)
  p_dry * abs(wheat_dry - exp_wheat) +
    p_wet * abs(wheat_wet - exp_wheat) +
    (1 - p_dry - p_wet) * abs(wheat_normal - exp_wheat)
  }

yield_mad_barley <- function(p_dry, p_wet) {
  exp_barley <- expected_yield_barley(p_dry, p_wet)
  p_dry * abs(barley_dry - exp_barley) +
    p_wet * abs(barley_wet - exp_barley) +
    (1 - p_dry - p_wet) * abs(barley_normal - exp_barley)
}

yield_mad_wheat(0.3, 0.2)
yield_mad_barley(0.3, 0.2)
```

@@

tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet) %>%
  gather(type, utility, dry_exp:wet_exp) %>%
  ggplot(aes(wheat_prop, utility)) +
 # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
#  geom_point(data = nash_eq, color = 'black', size = 3) +
 # geom_text(data = nash_eq, 
#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
#            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

```{r}
tibble(wheat_prop = seq(0, 1, .1),
       dry_exp = wheat_prop * wheat_dry + (1 - wheat_prop) * barley_dry,
       normal_exp = wheat_prop * wheat_normal + (1 - wheat_prop) * barley_normal,
      wet_exp = wheat_prop * wheat_wet + (1 - wheat_prop) * barley_wet,
      yield_exp = dry_exp + normal_exp + wet_exp, 
      lower_limit = pmin(dry_exp, normal_exp,wet_exp)) %>%
  ggplot(aes(yield_exp, lower_limit)) +
 # geom_hline(yintercept = nash_eq$utility, linetype = 2)+
#  geom_vline(xintercept = nash_eq$wheat_prop, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
#  geom_point(data = nash_eq, color = 'black', size = 3) +
 # geom_text(data = nash_eq, 
#            label = paste0('(', nash_eq$wheat_prop, ', ', nash_eq$utility, ')'),
#            nudge_x = 0.1, nudge_y = 0.05) +
  theme_classic()
```

@@

@@

sim_present <- brick('~/Downloads/b40.20th.track1.1deg.005.cam2.h0.PRECT.185001-200512.nc') %>%
  raster::extract(konya, df = TRUE) %>%
  select(-ID) %>%
  gather(time, rainfall) %>%
  mutate(rainfall = rainfall * 2.628e+9,
         year = rep(1850:2005, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
         month = rep(1:12, times = 156)) %>%
  select(-time) %>%
  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
  filter(between(water_year, 1932, 2005)) %>%
   filter(!(month %in% c(5:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(rainfall))

sim_past <- brick('~/Downloads/b40.mh6ka.1deg.003.cam2.h0.PRECT.080101-131612.nc') %>% 
  raster::extract(konya, df = TRUE) %>%
  select(-ID) %>%
  gather(time, rainfall) %>%
  mutate(rainfall = rainfall * 2.628e+9,
         year = rep(1:516, each = 12), # add hardcode dates instead of using ncdf time because of date shift with averaging period
         month = rep(1:12, times = 516)) %>%
  select(-time) %>%
  mutate(water_year = if_else(month %in% 10:12, year + 1L, year)) %>% 
   filter(!(month %in% c(5:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(rainfall))
```

```{r}
obs <- read_table('~/Downloads/pa17244.dat', 
           comment = "#", 
           col_names = c('year', 1:12), 
           na = c('-999.9', '-888.8')) %>%
  gather(month, precip, `1`:`12`) %>%
  group_by(month) %>%
  mutate(precip = if_else(is.na(precip), mean(precip, na.rm = TRUE), precip)) %>%
  ungroup %>%
  mutate(month = as.numeric(month),
         water_year = if_else(month %in% 10:12, year + 1, year)) %>% 
    filter(between(water_year, 1932, 2005)) %>%
 filter(!(month %in% c(5:9))) %>%
  group_by(water_year) %>%
  summarise(precip = sum(precip))
```

@@

tibble(droughts = (obs_past < 250)[1:500]) %>%
  mutate(time = rep(1:20, each = 25))%>%
  group_by(time) %>%
  summarise(test = sum(droughts)/25)
```

```{r}
simulation <- accumulate((obs_past < 250), ~learn(.x, .y), .init = agents) %>% 
  bind_rows(.id = 'time') %>%
  mutate(time = as.numeric(time)) %>%
  filter(time != 1)
```

so in this case drought risk is low enough all the time you never get any diversification

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

so we're imagining a yield of 1 for barley and 2 for wheat, with equal proportions of each. then in the next time step we'd revise the proportion to take away .25
```{r}
reallocate <- function(wheat_prop, risk) {
  wheat_prop + 1 * (1.18 * (1 - wheat_prop) - 1.6 * wheat_prop) * wheat_prop * (1 - wheat_prop)
}

@@

tibble(rainfall = (0:200) / 100,
       barley = calc_yield(rainfall, 'barley'),
       wheat = calc_yield(rainfall, 'wheat')) %>%
  gather(crop, yield, barley:wheat) %>%
  ggplot(aes(rainfall, yield, color = crop)) + 
  geom_line() +
  geom_hline(yintercept = yield_max_wheat * kcal_wheat) +
  geom_hline(yintercept = yield_max_barley * kcal_barley) +
  geom_hline(yintercept = 1016525 * 5, color ='red') + # cals for family of 5?
  geom_vline(xintercept = 0.3) +
  geom_vline(xintercept = .2)
```
